---
title: "Heatmap_bioxbrain"
author: "EmilyScott"
date: "2022-11-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library("Hmisc")

library(ggplot2)
library(hrbrthemes)
library(plotly)

library(ellipse)
library(RColorBrewer)

library(reshape2)
```

## R Correlation Heatmap: NB Behavior x CFOS Brain Density 

This is a heatmap containing the correlation between behavioral and brain region cfos data. The goal is to visualize the extent to which behavior and brain region are related.  

Import behavior data and brain cfos density data from the shared drive. 

## Import

```{r}
behxbrain <- read_csv("behxbrain.csv")
bb_df <- behxbrain
```

## Calculate Z Scores

```{r}
# mean and standard deviation (calculated inside z-score so not really necessary but nice to look at)
bb_mean <- colMeans(bb_df[sapply(bb_df, is.numeric)], na.rm = TRUE)
bb_sd <- sapply(bb_df[sapply(bb_df, is.numeric)], sd, na.rm = TRUE)

# z-score     z = (x - mean)/(sd)    (using only bb_df)
bb_z <- sapply(bb_df, function(bb_df)(bb_df-mean(bb_df, na.rm = T))/sd(bb_df,na.rm = T))
```

## Conduct Pearson Correlation

```{r}
#find correlation (r) and p-value
bb_z_cor <- rcorr(as.matrix(bb_z),  type = "pearson")

############### THIS MAY CHANGE WITH MORE MEASURES/REGIONS #############
#create table containing only correlations important : behavior vs brain
bb_z_cor_r <- bb_z_cor$r[1:11,12:15]
bb_z_cor_P <- bb_z_cor$P[1:11,12:15]
bb_z_cor_n <- bb_z_cor$n[1:11,12:15]

# other option :  cor(z[,1:11], z[,12:15], use="complete.obs", method="pearson")
```

## Plot Heatmap

```{r}
#prep r, p-value, stars for plotting
bb_melt_r<- melt(bb_z_cor_r)#melt correlation r value data for plotting
bb_melt_P<- melt(bb_z_cor_P)#melt correlation p value data for plotting
bb_melt_P$stars <- cut(bb_melt_P$value, breaks=c(-Inf, 0.001, 0.01, 0.05, Inf), label=c("***", "**", "*", "")) #p-value sig stars

bb_melt <- data.frame(BrainRegion = bb_melt_r$Var2, 
                      BehavioralMeasure = bb_melt_r$Var1,
                      r = bb_melt_r$value,
                      P = bb_melt_P$value,
                      stars = bb_melt_P$stars )
#save data
write.csv(bb_melt, "bb_melt.csv")

#plot and save 
tiff("bb_heatmap.tiff", height = 30, width = 20, units = 'cm', compression = "lzw", res = 300)

ggplot(data = bb_melt, aes(x=BrainRegion, y=BehavioralMeasure, fill=r)) +
  geom_tile() +
   geom_text(aes(label=stars), color="black", size=5)+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation")

dev.off()
```
